---
title: "ISyE 6414 Project 2"
author: "Xiaochen Yan"
date: "2023-11-23"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("/Users/lazayxc/Desktop/GT/Fall 23/ISYE 6414 Regression/6414 project/analysis files/2. model")

```

We eliminated the non-positive values of the response variables (Dividend Payout Ratio)
```{r,warning=FALSE,message=FALSE}
data1<-read.csv("pos_data.csv",sep=',',header=TRUE)[-1]
dim(data1)

```

----------Part 3: Data Description------------

3.1 Descriptive statistics

The maximum values of the five variables show that there are some unusual observations. 

```{r}
summary(data1)
```
3.2 Scatter Plot

The scatter plot also shows unusual observations. Besides, due to unusual values, the relationship between response variable and independent variables is not clear.

```{r}
plot(data1[,-ncol(data1)-1])
```
3.3 Process Unusual Observations

```{r,warning=FALSE,message=FALSE}

# Replace values greater than the 99% quantile with the 99% quantile value, excluding "Industry"
data_99 <- data1  # Copy the original data frame

quantile_99 <- apply(data_99, 2, function(x) quantile(x, probs = 0.99))
data_99["Profitability"] <- apply(data_99["Profitability"], 2, function(x, q) ifelse(x > q, q, x), quantile_99)
data_99["Cash.Ratio"] <- apply(data_99["Cash.Ratio"], 2, function(x, q) ifelse(x > q, q, x), quantile_99)
data_99["Sales.Growth"] <- apply(data_99["Sales.Growth"], 2, function(x, q) ifelse(x > q, q, x), quantile_99)
data_99["Debt.to.Equity.Ratio"] <- apply(data_99["Debt.to.Equity.Ratio"], 2, function(x, q) ifelse(x > q, q, x), quantile_99)
data_99["Dividend.Payout.Ratio"] <- apply(data_99["Dividend.Payout.Ratio"], 2, function(x, q) ifelse(x > q, q, x), quantile_99)

# Print the modified data frame
summary(data_99)
```
3.4 Data Histogram 

```{r}
hist(data_99$Dividend.Payout.Ratio,main="",xlab="Count of Dividend Payout Ratio", border="gold",col="darkblue")
hist(data_99$Profitability,main="",xlab="Count of Profitability", border="gold",col="darkblue")
hist(data_99$Cash.Ratio,main="",xlab="Count of Profitability", border="gold",col="darkblue")
hist(data_99$Sales.Growth,main="",xlab="Count of Profitability", border="gold",col="darkblue")
hist(data_99$Debt.to.Equity.Ratio,main="",xlab="Count of Profitability", border="gold",col="darkblue")
```
----------Part 4: Analysis----------------

4.1 Split data set and correlation analysis

```{r,message=FALSE,warning=FALSE}
set.seed(1234)

sample_size = floor(0.8*nrow(data_99))
picked = sample(seq_len(nrow(data_99)),size = sample_size)

data = data_99[picked,]
y <- data$Dividend.Payout.Ratio

attach(data)

```


```{r}
summary(y)
print(length(y))
```
Since the response value is a ratio, we choose the majority of data (less than 2) for histogram. 

```{r}
adj_y = y[y<2]
length(adj_y)
hist(adj_y,main="",xlab="Count of Dividend Payout Ratio",breaks = 10,xlim = c(0, 2), border="gold",col="darkblue")
```
```{r}
corr = cor(data)
print(corr)
```



4.2 Box-Cox Transformation

Apply the box-cox and the result shows that we need to apply natural logarithm transformation to the response variable
```{r}
library(MASS)
result <- boxcox(y ~ 1, data = data)
result
```
4.3 First-order model
4.3.1 Regression

F-statistic shows the model as a whole is significant. The coefficient of Sales.Growth and Debt.to.Equity.Ratio is significant.

### First -order model

```{r}

model1 = lm(log(y)~Profitability+Cash.Ratio+Sales.Growth+Debt.to.Equity.Ratio+factor(INDUSTRY),data = data)
summary(model1)

```

test for subset of coefficient
```{r}
reduced_model1 = lm(log(y) ~ Cash.Ratio + Sales.Growth,data=data)
anova(reduced_model1, model1)

```

4.3.2 Model Diagnostics
Analysis of residuals:

```{r,message=FALSE,warning=FALSE}
res = stdres(model1)


par(mfrow = c(2,2))
plot(model1$fitted.values, res, xlab = "Fitted Values", ylab = "Residuals",pch = 19)
abline(h = 0)

# plot(Profitability, res, xlab = "Profitability", ylab = "Residuals",xlim=c(0,1), pch = 19)
# abline(h = 0)
# plot(Cash.Ratio, res, xlab = "Cash.Ratio", ylab = "Residuals",xlim=c(0,5), pch = 19)
# abline(h = 0)
# plot(Sales.Growth, res, xlab = "Sales.Growth", ylab = "Residuals",xlim=c(0,3), pch = 19)
# abline(h = 0)
# plot(Debt.to.Equity.Ratio, res, xlab = "Debt.to.Equity.Ratio", ylab = "Residuals",xlim=c(0,50), pch = 19)
# abline(h = 0)
# plot(INDUSTRY, res, xlab = "INDUSTRY", ylab = "Residuals", pch = 19)
# abline(h = 0)


hist(res, xlab="Residuals", main= "Histogram of Residuals")

qqnorm(res,ylim=c(-5,5))
qqline(res)

cook = cooks.distance(model1)
plot(cook,type="h",lwd=3, ylab = "Cook’s Distance")

```
Use Cook's distance to check if there are any outliers.
```{r}
cook = cooks.distance(model1)
plot(cook,type="h",lwd=3, ylab = "Cook’s Distance")
hurdle = 4/length(data)

indices <- which(cook > hurdle)
if(length(indices) > 0) {
    print(cook[indices])
} else {
    print("No Outliers")
}

```
Multicollinearity:

```{r,warning=FALSE,message=FALSE}
library(car)

vif(model1)
max(10,1/(1-summary(model1)$r.squared))
```

Apply transformation to the predict variables. Only Debt.to.Equity.Ratio is positive so we will use log transformation to it.
```{r}
model1_trans = lm(log(y)~Profitability+sqrt(Cash.Ratio)+Sales.Growth+log(Debt.to.Equity.Ratio)+factor(INDUSTRY),data = data)
summary(model1_trans)
```




```{r,warning=FALSE,message=FALSE}

res = stdres(model1_trans)
cook = cooks.distance(model1_trans)

par(mfrow = c(3,3))
plot(model1_trans$fitted.values, res, xlab = "Dividend.Payout.Ratio", ylab = "Residuals",pch = 19)
abline(h = 0)
plot(Profitability, res, xlab = "Profitability", ylab = "Residuals",xlim=c(0,1), pch = 19)
abline(h = 0)
plot(sqrt(Cash.Ratio), res, xlab = "square root of Cash.Ratio", ylab = "Residuals",xlim=c(0,5), pch = 19)
abline(h = 0)
plot(Sales.Growth, res, xlab = "Sales.Growth", ylab = "Residuals",xlim=c(0,3), pch = 19)
abline(h = 0)
plot(log(Debt.to.Equity.Ratio), res, xlab = "log of Debt.to.Equity.Ratio", ylab = "Residuals",xlim=c(0,5), pch = 19)
abline(h = 0)
plot(INDUSTRY, res, xlab = "INDUSTRY", ylab = "Residuals", pch = 19)
abline(h = 0)


hist(res, xlab="Residuals", main= "Histogram of Residuals")

qqnorm(res,ylim=c(-5,5))
qqline(res)

plot(cook,type="h",lwd=3, ylab = "Cook’s Distance")


```


```{r}

vif(model1_trans)
max(10,1/(1-summary(model1_trans)$r.squared))

```

## Poisson
```{r}
model3 = glm(round(y*100) ~ Profitability+Cash.Ratio+Sales.Growth+Debt.to.Equity.Ratio+factor(INDUSTRY),data=data, family="poisson" )

summary(model3)
```
Test for overall regression

```{r}
1-pchisq((213187- 144853), 1918-1892)
```
Goodness of Fit Test - Deviance Test

```{r}
with(model3, cbind(res.deviance = deviance, df = df.residual,
                           p = 1- pchisq(deviance, df.residual)))
```
p-value = 0. Reject the null hypothesis of good fit. 

Residual Analysis

```{r}
res <- resid(model3,type="deviance")

par(mfrow = c(3,3))
plot(model3$fitted.values, res, xlab = "Dividend.Payout.Ratio", ylab = "Residuals", pch = 19)
abline(h = 0)
plot(Profitability, res, xlab = "Profitability", ylab = "Residuals",xlim=c(0,1), pch = 19)
abline(h = 0)
plot(Cash.Ratio, res, xlab = "pk.Ratio", ylab = "Residuals",xlim=c(0,5), pch = 19)
abline(h = 0)
plot(Sales.Growth, res, xlab = "Sales.Growth", ylab = "Residuals",xlim=c(0,3), pch = 19)
abline(h = 0)
plot(Debt.to.Equity.Ratio, res, xlab = "Debt.to.Equity.Ratio", ylab = "Residuals",xlim=c(0,50), pch = 19)
abline(h = 0)
plot(INDUSTRY, res, xlab = "INDUSTRY", ylab = "Residuals", pch = 19)
abline(h = 0)


hist(res, xlab="Residuals", main= "Histogram of Residuals")

qqnorm(res)
qqline(res)

plot(cook,type="h",lwd=3, ylab = "Cook’s Distance")

```

```{r}
## Do we have overdispersion?
wdf <- model3$df.residual # n-p-1
dev <- model3$deviance
overdisp <- dev/wdf
print(overdisp)
```
```{r}
## Assuming GOF: Deviance ~ N(0,1)
##  Evaluate Deviance residuals vs the Normal quantile with prob=0.99995
dev_residuals <- residuals(model3, type = "deviance")
outliers <- which(abs(dev_residuals) > qnorm(.99995))
length(outliers)
```
959 outliers, indicating overdispersion.


Apply an overdispersed Poisson Regression
```{r}
quasipoisson_model <- glm(round(y*100) ~ Profitability+Cash.Ratio+Sales.Growth+Debt.to.Equity.Ratio+factor(INDUSTRY), data=data, family="quasipoisson")
summary(quasipoisson_model)

```





### Second -order model

```{r}
model2.1 = lm(log(y)~(Profitability+sqrt(Cash.Ratio)+Sales.Growth+log(Debt.to.Equity.Ratio))**2+factor(INDUSTRY)+I(Profitability^2)+I(sqrt(Cash.Ratio)^2)+I(Sales.Growth^2)+I(log(Debt.to.Equity.Ratio^2)),data=data)
summary(model2.1)

```

```{r}
model2.2 = lm(log(y)~(Profitability+Cash.Ratio+Sales.Growth+log(Debt.to.Equity.Ratio))**2+factor(INDUSTRY)+I(Profitability^2)+I(Cash.Ratio^2)+I(Sales.Growth^2)+I(log(Debt.to.Equity.Ratio^2)),data=data)
summary(model2.2)

```

```{r,message=FALSE,warning=FALSE}
res = stdres(model2.1)
cook = cooks.distance(model2.1)

par(mfrow = c(3,3))
plot(model2.1$fitted.values, res, xlab = "Dividend.Payout.Ratio", ylab = "Residuals", pch = 19)
abline(h = 0)
plot(Profitability, res, xlab = "Profitability", ylab = "Residuals",xlim=c(0,1), pch = 19)
abline(h = 0)
plot(Cash.Ratio, res, xlab = "pk.Ratio", ylab = "Residuals",xlim=c(0,5), pch = 19)
abline(h = 0)
plot(Sales.Growth, res, xlab = "Sales.Growth", ylab = "Residuals",xlim=c(0,3), pch = 19)
abline(h = 0)
plot(Debt.to.Equity.Ratio, res, xlab = "Debt.to.Equity.Ratio", ylab = "Residuals",xlim=c(0,50), pch = 19)
abline(h = 0)
plot(INDUSTRY, res, xlab = "INDUSTRY", ylab = "Residuals", pch = 19)
abline(h = 0)


hist(res, xlab="Residuals", main= "Histogram of Residuals")

qqnorm(res,ylim=c(-5,5))
qqline(res)

plot(cook,type="h",lwd=3, ylab = "Cook’s Distance")


```

```{r}
model2 = lm(log(y)~(Profitability+sqrt(Cash.Ratio)+Sales.Growth+log(Debt.to.Equity.Ratio))**2+factor(INDUSTRY),data=data)
summary(model2)

```


```{r}
model2_adj = lm(log(y)~Profitability+sqrt(Cash.Ratio)+Sales.Growth+log(Debt.to.Equity.Ratio)+factor(INDUSTRY)+Profitability*sqrt(Cash.Ratio)+Profitability*Sales.Growth+Profitability*log(Debt.to.Equity.Ratio)+sqrt(Cash.Ratio)*log(Debt.to.Equity.Ratio) ,data = data)
summary(model2_adj)
anova(model2_adj, model2)

```
We fail to reject Null hypothesis! so we can choose the model with subset of coefficients


```{r,message=FALSE,warning=FALSE}
res = stdres(model2_adj)
cook = cooks.distance(model2_adj)

par(mfrow = c(3,3))
plot(model2_adj$fitted.values, res, xlab = "Dividend.Payout.Ratio", ylab = "Residuals", pch = 19)
abline(h = 0)
plot(Profitability, res, xlab = "Profitability", ylab = "Residuals",xlim=c(0,1), pch = 19)
abline(h = 0)
plot(Cash.Ratio, res, xlab = "pk.Ratio", ylab = "Residuals",xlim=c(0,5), pch = 19)
abline(h = 0)
plot(Sales.Growth, res, xlab = "Sales.Growth", ylab = "Residuals",xlim=c(0,3), pch = 19)
abline(h = 0)
plot(Debt.to.Equity.Ratio, res, xlab = "Debt.to.Equity.Ratio", ylab = "Residuals",xlim=c(0,50), pch = 19)
abline(h = 0)
plot(INDUSTRY, res, xlab = "INDUSTRY", ylab = "Residuals", pch = 19)
abline(h = 0)


hist(res, xlab="Residuals", main= "Histogram of Residuals")

qqnorm(res,ylim=c(-5,5))
qqline(res)

plot(cook,type="h",lwd=3, ylab = "Cook’s Distance")


```
```{r}
outliers = which(cook>1)
length(outliers)
data_without_outliers <- data[-outliers, ]
y_without_outliers <- y[-outliers, ]
```


```{r}
model2_no_outlier = lm(log(y_without_outliers)~Profitability+Cash.Ratio+Sales.Growth+log(Debt.to.Equity.Ratio)+factor(INDUSTRY)+Profitability*Cash.Ratio+Profitability*Sales.Growth+Profitability*log(Debt.to.Equity.Ratio)+Cash.Ratio*log(Debt.to.Equity.Ratio) ,data = data_without_outliers)
summary(model2_no_outlier)

```

```{r,message=FALSE,warning=FALSE}
res = stdres(model2_no_outlier)
cook = cooks.distance(model2_no_outlier)

par(mfrow = c(3,3))
plot(model2_no_outlier$fitted.values, res, xlab = "Dividend.Payout.Ratio", ylab = "Residuals", pch = 19)
abline(h = 0)
plot(Profitability[-outliers, ], res, xlab = "Profitability", ylab = "Residuals",xlim=c(0,1), pch = 19)
abline(h = 0)
plot(Cash.Ratio[-outliers, ], res, xlab = "pk.Ratio", ylab = "Residuals",xlim=c(0,5), pch = 19)
abline(h = 0)
plot(Sales.Growth[-outliers, ], res, xlab = "Sales.Growth", ylab = "Residuals",xlim=c(0,3), pch = 19)
abline(h = 0)
plot(Debt.to.Equity.Ratio[-outliers, ], res, xlab = "Debt.to.Equity.Ratio", ylab = "Residuals",xlim=c(0,50), pch = 19)
abline(h = 0)
plot(INDUSTRY[-outliers], res, xlab = "INDUSTRY", ylab = "Residuals", pch = 19)
abline(h = 0)


hist(res, xlab="Residuals", main= "Histogram of Residuals")

qqnorm(res,ylim=c(-5,5))
qqline(res)

plot(cook,type="h",lwd=3, ylab = "Cook’s Distance")


```

4.5 Model Selection

### Model selection
```{r}
detach(data)

test=data_99[-picked,]

```


```{r}
pred1 <- predict(model1_trans, test, interval = 'prediction')
test.pred1 <- pred1[,1]
test.lwr1 <- pred1[,2]
test.upr1 <- pred1[,3]
# Mean Squared Prediction Error (MSPE)
mean((test.pred1-test$Dividend.Payout.Ratio)^2)

# Mean Absolute Prediction Error (MAE)
mean(abs(test.pred1-test$Dividend.Payout.Ratio))

# Mean Absolute Percentage Error (MAPE)
mean(abs(test.pred1-test$Dividend.Payout.Ratio)/test$Dividend.Payout.Ratio)

# Precision Measure (PM)
sum((test.pred1-test$Dividend.Payout.Ratio)^2)/sum((test$Dividend.Payout.Ratio-mean(test$Dividend.Payout.Ratio))^2)

# CI Measure (CIM)
(sum(test$Dividend.Payout.Ratio<test.lwr1)+sum(test$Dividend.Payout.Ratio>test.upr1))/nrow(test)


```


```{r}

pred1 <- predict(model2_adj, test, interval = 'prediction')
test.pred1 <- pred1[,1]
test.lwr1 <- pred1[,2]
test.upr1 <- pred1[,3]
# Mean Squared Prediction Error (MSPE)
mean((test.pred1-test$Dividend.Payout.Ratio)^2)

# Mean Absolute Prediction Error (MAE)
mean(abs(test.pred1-test$Dividend.Payout.Ratio))

# Mean Absolute Percentage Error (MAPE)
mean(abs(test.pred1-test$Dividend.Payout.Ratio)/test$Dividend.Payout.Ratio)

# Precision Measure (PM)
sum((test.pred1-test$Dividend.Payout.Ratio)^2)/sum((test$Dividend.Payout.Ratio-mean(test$Dividend.Payout.Ratio))^2)

# CI Measure (CIM)
(sum(test$Dividend.Payout.Ratio<test.lwr1)+sum(test$Dividend.Payout.Ratio>test.upr1))/nrow(test)

```

```{r}

pred1 <- predict(model2_no_outlier, test, interval = 'prediction')
test.pred1 <- pred1[,1]
test.lwr1 <- pred1[,2]
test.upr1 <- pred1[,3]
# Mean Squared Prediction Error (MSPE)
mean((test.pred1-test$Dividend.Payout.Ratio)^2)

# Mean Absolute Prediction Error (MAE)
mean(abs(test.pred1-test$Dividend.Payout.Ratio))

# Mean Absolute Percentage Error (MAPE)
mean(abs(test.pred1-test$Dividend.Payout.Ratio)/test$Dividend.Payout.Ratio)

# Precision Measure (PM)
sum((test.pred1-test$Dividend.Payout.Ratio)^2)/sum((test$Dividend.Payout.Ratio-mean(test$Dividend.Payout.Ratio))^2)

# CI Measure (CIM)
(sum(test$Dividend.Payout.Ratio<test.lwr1)+sum(test$Dividend.Payout.Ratio>test.upr1))/nrow(test)

```